Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-08-07 other/confnoncoherent/paper.tex
added to convergence proof. Getting there, but needs some cleanup.

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.  Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
\end{equation}
and define the continuous function
\[
G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\] 
If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

>>>>

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.  Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
\end{equation}
and define the continuous function
\[
G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\] 
If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

### 2012-08-07 paper.tex
added to convergence proof. Getting there, but needs some cleanup.

\section{Mackentun's algorithm}\label{sec:least-squar-estim}

@@

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$. %where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  
Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
\]
\[
h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

@@

@@

\begin{lemma}\label{lem:convtoexpGlamL} 
$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
\[
0 \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)},
\]
and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
\[ 
0 \leq \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)}.
\]
Thus,
\begin{align*}
0 &\leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} \\ 
& \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} + \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)} \\
& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)| \\
&\leq 2\sup_{x \in [-\pi, \pi)} \sabs{G(x) - G_L(x)},
\end{align*}
and the last line converges amost surely to zero by Lemma~\ref{lem:uniflawGGL}.
\end{IEEEproof}

@@

\begin{lemma}\label{lem:uniflawGGL}
The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
\[
\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - G(x)} \rightarrow 0
\] 
almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put $T(x) = \expect G_L(x)$ and observe that
\begin{align*}
T(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
\end{align*}
where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, then $\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.  Now
\begin{align*}
\sup_{\lambda \in [-\pi, \pi)}&\sabs{G_L(\lambda) - G(\lambda)} \\
&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}
\end{align*}
and Lemma~\ref{lem:uniflawTGL} shows that $\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.
\end{IEEEproof}

\begin{lemma}\label{lem:uniflawTGL} 
Put $T(x) = \expect G_L(x)$, then
\[
\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - T(x)} \rightarrow 0
\] 
almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
BLERG
\end{IEEEproof}

>>>>

\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}

@@

\begin{theorem}\label{thm:consistency} (Almost sure convergence)
Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.   Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$. %where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.  
Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
\[
\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
\] 
Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
\begin{equation}\label{eq:RiandPhii}
R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
\end{equation}
and define the continuous function
\[
G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
\]
\[
h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
\]
If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
\begin{enumerate}
\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
\end{enumerate}
\end{theorem}

@@

The assumption that $\expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.  This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.  We will also use that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}

@@

\begin{lemma}\label{lem:convtoexpGlamL} 
$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
\[
0 \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)},
\]
and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
\[ 
0 \leq \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)}.
\]
Thus,
\begin{align*}
0 &\leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} \\ 
& \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} + \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)} \\
& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)| \\
&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G(\lambda) - G_L(\lambda)},
\end{align*}
and the last line converges amost surely to zero by Lemma~\ref{lem:uniflawGGL}.
\end{IEEEproof}

@@

\begin{lemma}\label{lem:uniflawGGL}
The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
\[
\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0
\] 
almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put $T(\lambda) = \expect G_L(\lambda)$ and write
\begin{align*}
&\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}.
\end{align*}
Now,
\begin{align*}
T(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
\end{align*}
where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, it follows that 
\[
\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0
\]  
almost surely as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
\[
\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
\]
almost surely as $L\rightarrow \infty$.
\end{IEEEproof}

\begin{lemma}\label{lem:uniflawTGL} 
Put $T(\lambda) = \expect G_L(\lambda)$, then
\[
\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
\] 
almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put $D_L(\lambda) = G_L(\lambda) - T(\lambda)$ and let
\[
\lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
\]
be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n$ be the interval $[\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
\begin{align*}
\sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
&= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
&\leq A_L + B_L,
\end{align*}
where 
\[
A_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
\]
\[
B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
\]
Lemma~\ref{lem:supDLlambdan} shows that for any finite $N$, and any $\epsilon > 0$, 
\[
\prob\left\{ \lim_{L\rightarrow\infty} A_L > \epsilon \right\} = 0,
\]
i.e., $A_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
\[
\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
\] 
Choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ and
\begin{align*}
\prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
&\leq \prob\left\{\lim_{L\rightarrow\infty}(A_L + B_L) > 3\epsilon \right\} \\
&\leq \prob\left\{\lim_{L\rightarrow\infty} (A_L + B_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
&\leq \prob\left\{\lim_{L\rightarrow\infty} A_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
&= 0.
\end{align*}
Thus, $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
\end{IEEEproof}

\begin{lemma}\label{lem:supDLlambdan}
If $N$ is finite then $A_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put
\[
Z_i(\lambda) = \begin{cases}
R_i\cos(\lambda + \Phi_i) , & i \in P \\
R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
\end{cases}
\]
so that
\[
D_L(\lambda) = G_L(\lambda) - T(\lambda) = \frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
\]
Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
\[
\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
\]
almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Now, since $N$ is finite,
\[
A_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
\]
almost surely to zero as $L \rightarrow \infty$.
\end{IEEEproof}

\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
\[
\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
\]
\end{lemma}
\begin{IEEEproof}
Observe that
\begin{align*}
&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
&\leq  \abs{G_L(\lambda) - \expect G_L(\lambda) - G_L(\lambda_n) + \expect G_L(\lambda_n)} \\
&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
\end{align*}
the last line following from Jensens inequality.  Put
\[
C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) }
\]
then
\[
B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L.
\]
because 
\[
 \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect \abs{G_L(\lambda) - G_L(\lambda_n)} \leq \expect C_L
\]
follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and also that
\[
\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
\]
So,
\begin{align*}
\prob&\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
\end{align*}
\end{IEEEproof}

\begin{lemma}\label{lem:CL}
$\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and for any $\epsilon > 0$,
\[
\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
\]
\end{lemma}
\begin{IEEEproof}
If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
\[
\abs{\cos(\lambda) - \cos(\lambda_n)} \leq \tfrac{2\pi}{N} \;\;\; \text{and} \;\;\; \abs{\cos\sfracpart{\lambda} - \cos\sfracpart{\lambda_n}} \leq \tfrac{2\pi}{N}.
\]
Because this results does not depend on $n$,
\[
\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda)} \leq R_i \frac{2\pi}{N}
\]
for all $i = 1, \dots, L$.  Now,
\begin{align*}
C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i=1}^L Z_i(\lambda) - Z_i(\lambda_n)} \\
&\leq \frac{1}{L}\sum_{i=1}^L \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
&\leq \frac{2\pi}{N L}\sum_{i=1}^L R_i \rightarrow \frac{2\pi}{N}\expect R_i
\end{align*}
almost surely to $\frac{2\pi}{N}\expect R_i$ as $L \rightarrow\infty$ by the strong law of large numbers, and so
\begin{align*}
\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} \\
&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i=1}^L R_i > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
\end{align*}
Also $\expect C_L \leq \frac{2\pi}{N}\expect R_i$.
\end{IEEEproof}

\begin{lemma}\label{lem:coslipshitz}
Let $x$ and $\delta$ be real numbers.  Then
\[
\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta},
\]
and 
\[
\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
\]
\end{lemma}
